---
title: "MATH 216 Homework 4"
author: "Delaney Moran"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# For data manipulation and visualization
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
# For US county and state maps
suppressPackageStartupMessages(library(maps))
# For loading in shapefiles
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(maptools))
# For interactive maps
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(spdep))
```

## Admistrative:

Please indicate

* Who you collaborated with: Phil, Kyler, Christian, Aminata
* Roughly how much time you spent on this HW: 7
* What gave you the most trouble:
* Any comments you have: 



## Question 1:

```{r, cache=TRUE, echo=FALSE}
# Load state and county map of US in 2010 from the maps package and convert them
# to data frames so that we can ggplot them.
US_state <- map_data("state") %>% 
  tbl_df()
US_county <- map_data("county") %>% 
  tbl_df()

clean_text <- function(text){
  text <- gsub("[^[:alnum:]]", "", text)
  text <- gsub(" ", "", text)
  text <- tolower(text)
  return(text)
}

COUNTY <- read.csv("~/Documents/HW-4/COUNTY.csv")

COUNTY1 <- COUNTY %>% 
  mutate(COUNTY = clean_text(COUNTY),
         STATE = clean_text(STATE)) %>% 
  select(CID, STATE, COUNTY, PBUSH) %>%  
  mutate(PBUSH = as.numeric(as.character(PBUSH)))

US_county1 <- US_county %>% 
    mutate(COUNTY = clean_text(subregion),
           STATE = clean_text(region)) %>% 
    select(-subregion, -region)

US_county2 <- left_join(US_county1, COUNTY1, by = c("COUNTY", "STATE"))

ggplot(data=NULL) +
  geom_polygon(data=US_county2, aes(x=long, y=lat, group=group, fill = PBUSH)) +
  scale_fill_gradient2(name = "", low="blue", mid="white", high="red", midpoint = 0.5) +
  geom_path(data=US_county, aes(x=long, y=lat, group=group), col="black", size=0.05) +
  geom_path(data=US_state, aes(x=long, y=lat, group=group), col="black", size=0.15) +
  coord_map() +
  ggtitle("2000 Election Voting Behavior by County") +
  ylab("Latitude") +
  xlab("Longitude")
    

```


### Which states exhibit the greatest **within state heterogeneity** in voting?
Come up with a mathematical justification.

```{r, cache=FALSE, echo=FALSE}

Homogeneity <- US_county2 %>% 
  distinct(COUNTY) %>% 
  group_by(STATE) %>% 
  summarise(SD = sd(PBUSH, na.rm = TRUE)) %>% 
  arrange(-SD)

knitr::kable(head(Homogeneity, 5), caption = "Most Heterogeneous States")



```

The states listed above have the largest standard deviations in percent voting for Bush.



## Question 2:

In this question, please make 4 separate Leaflet maps showing a choropleth map for each ethnic group.

* White
* African-American
* Hispanic
* Asian and Pacific Island
* Other, including those of two or more races


```{r, cache=FALSE, echo=FALSE}
vt_race_data <- read.csv("~/Documents/HW-4/vt_race_data.csv")

vt_race <- vt_race_data %>% 
  select(Geo_FIPS, Geo_NAME, Geo_COUNTY, Geo_TRACT, SE_T014_001, SE_T014_003, SE_T014_004, 
         SE_T014_005, SE_T014_006, SE_T014_007, SE_T014_008, SE_T014_009, SE_T014_010) %>% 
  rename(Total_pop = SE_T014_001, 
         White = SE_T014_003,
         Black = SE_T014_004,
         American_Indian = SE_T014_005,
         Asian = SE_T014_006,
         Hawaiian_Islander = SE_T014_007,
         Other = SE_T014_008,
         Two_or_more = SE_T014_009,
         Hispanic = SE_T014_010) %>% 
  mutate(Other_incl = American_Indian + Hawaiian_Islander + Other + Two_or_more) %>% 
  select(-American_Indian, -Hawaiian_Islander, -Other, -Two_or_more)

census_race <- vt_race %>% 
  mutate(prop_white = White/Total_pop, 
         prop_black = Black/Total_pop, 
         prop_asian = Asian/Total_pop,
         prop_hisp = Hispanic/Total_pop,
         prop_other = Other_incl/Total_pop)

shapefile_name <- paste(getwd(), "/VT_census_tracts/tl_2015_50_tract.shp", sep="")
VT <- readOGR(shapefile_name, layer = "tl_2015_50_tract", verbose = FALSE)


pal_white<- colorNumeric(palette = "Purples", domain = census_race$prop_white)
pal_hisp<- colorNumeric(palette = "OrRd", domain = census_race$prop_hisp)
pal_black<- colorNumeric(palette = "Blues", domain = census_race$prop_black)
pal_asian<- colorNumeric(palette = "Greens", domain = census_race$prop_asian)



white <- leaflet(VT) %>%
  addTiles() %>% 
  addPolylines(color="black", weight=1) %>% 
  addPolygons(stroke = FALSE, fillOpacity = 1, 
              color = ~pal_white(census_race$prop_white)) %>% 
  addLegend("bottomright", pal = pal_white, values = ~census_race$prop_white, 
            title = "Proportion White")
white
  
  
hisp <- leaflet(VT) %>%
  addTiles() %>% 
  addPolylines(color="black", weight=1) %>% 
  addPolygons(stroke = FALSE, fillOpacity = 1, 
              color = ~pal_hisp(census_race$prop_hisp)) %>% 
  addLegend("bottomright", pal = pal_hisp, values = ~census_race$prop_hisp, 
            title = "Proportion Hispanic")
hisp

black <- leaflet(VT) %>%
  addTiles() %>% 
  addPolylines(color="black", weight=1) %>% 
  addPolygons(stroke = FALSE, fillOpacity = 1, 
              color = ~pal_black(census_race$prop_black)) %>% 
  addLegend("bottomright", pal = pal_black, values = ~census_race$prop_black, 
            title = "Proportion Black")
black

asian <- leaflet(VT) %>%
  addTiles() %>% 
  addPolylines(color="black", weight=1) %>% 
  addPolygons(stroke = FALSE, fillOpacity = 1, 
              color = ~pal_asian(census_race$prop_asian)) %>% 
  addLegend("bottomright", pal = pal_asian, values = ~census_race$prop_asian, 
            title = "Proportion Asian")
asian


```


### Write-Up

Comment on general ethnic demographic trends.

There appears to be a region in the North West corner of VT with very high levels of Asians and African Americans. Additionally, there are many isolated clusters of Hispanic populations throughout the state. Interestingly, there is a region in the center of the state where several census tracts with very high proportions of Hispanics surround a very white area that appears to be Rutland. 

